How to Recognize Useless ``Bad Smells''

### 2016-03-03 Manuscript.tex
Updates from ShareLaTeX

\title{Learning Which Bad Smells are 
Real Stinkers?}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract}

\end{abstract}

\keywords{Defect prediction; Configuration; Planning; Decision trees.}
\section{Introduction}

@@

\input{tex/jur_char.tex}

The thesis of this paper is that the XTREE tool can stop developers
wasting time on useless software refactorings. Before presenting experiments
that supports that claim, this section motivates the need for 
such tools.

In many fields,  practitioners hang on to persistent beliefs, even
if the evidence for those beliefs is not strong. For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
There is growing evidence that amongst programmers,
many of their  widely-held beliefs are inaccurate and misleading. 
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer two
specific examples
were  widely-held views in SE that are   now questionable, given recent results. 
Firstly, Nagappan et al. analyzed~\cite{mei15} 
how the {\tt goto} statement was used  in 11,000 Github repositories, 
they found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.

Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. Boehm and Basili comment that this phase delay effect
``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
After reading the literature, and surveying academics
and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
this effect seems now heavily mitigated by $21^{st}$ software development languages,
tools and development practices.  When we mined the 171 software projects shepherded
by the SEI in the period 2006-2014, we found that phase delay results
in only minor changes in the 
time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.

These two studies are symptomatic
of a widespread malaise in software engineering.
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  `` (a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

One area where poorly-justifed beliefs can waste valuable project resources is
refactoring in order to remove code bad smells.

@@

>>>>

\title{ How to Recognize Useless ``Bad Smells''}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,~xipengshen\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract}
Developers, text books, and tools
disagree on what bad smells should be refactored.
While it useful to fix  some bad smells, others are
 can be ignored.
How to distinguish useful from useless bad smells?

We propose RANK, a tool that judges how badly
a project needs a particular   refactoring.
RANK is tuned using the historical data of the current project.
Using RANK, we   show that the range of truly bad
``bad smells'' changes from project to project. Hence,
before refactoring, it is important to run a critique tool
like RANK to stop work on useless refactorings.
Further, using RANK, it is possible to show that many
of the bad smells endorsed by standard texts or tools
are not useful for particular projects. 
Based on these results, we endorse  
guiding refactoring via bad smells. However,  
it is vital to RANK the bad smells defined for a 
particular project.

\vspace{3mm}
\noindent
{\bf Categories/Subject Descriptors:} D.2 [Software Engineering]; I.2 [Artificial Intelligence];

\noindent
{\bf Keywords:} Bad smells, refactoring,
performance prediction,  decision trees 
\end{abstract}

\section{Introduction}
\begin{raggedleft}
{\em My offense is rank, it smells to heaven.}\\
-- Hamlet, Act 3, Scene 3

\end{raggedleft}

~

% Developers, tools, and research papers are either vague or disagree
% about what code refactorings are most important to perform, or ignore.
% As noted by Shantnawi~\cite{Shatnawi10} there is a dearth of
% studies that formulate 
% precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
% claim that OO design metrics (e.g. depth of inheritance,
% classes per method, etc) can be used to predict
% code quality. Yet they their paper never comments at what point
% a particular code measurement turns from ``ok'' to ``bad smell''.
% Similary, code smell detection tools work for a certain
% range of

% When there are many possible code refactorings, it becomes difficult
% to choose which ones to execute. 
% Precious project resources can
% be wasted if developers spend too much time on useless refactorings.

According to   Fowler~\cite{fowler99}, code smells (a.k.a. bad smells)
are ``a surface indication that usually corresponds to a deeper problem''.
Much research enthusiastically endorses ``bad smell''-based refactoring, as witnessed
by the  
recent literature of Tufano et al. at ASE'15~\cite{Tufano2015} which
lists dozens of  
papers proposing bad smell detection and repair tools.

Should we remove bad smells whenever we find them? Perhaps not.

Firstly, as noted by Shantnawi~\cite{Shatnawi10} there are  there are few studies that formulate 
precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment at what point
a particular code measurement turns from ``ok'' to ``bad smell''.

Secondly, it is not clear that all code smells are relevant to
all projects.
Developers often disagree on what bad smells, if any at all,
should be refactored.
For example, 
Mantyla et al.~\cite{Mantyla2004} found that within an organization,
 developers had different perceptions on the importance of different
 bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
 surveyed 85 professional developers to conclude that it remains
 open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
 Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
 performing maintenance tasks that took three to four weeks to complete. 
 After modifying nearly 300 Java files, it was observed
 that none of the dozen investigated code smells explored in 
 that study were mostly irrelevant to to that maintenance work.

 How do we reconcile the enthusiasm of Fower,  Tufano, Darcy, Kemerer et al.
 with the more pessimism of Mantyla, Yamashita, 
 and Sj\"oberg et al.? The starting
 point of this paper is a comment
 from Mantyla et al.~\cite{Mantyla2004}. They say that
 when making  refactoring decisions, it is not enough to
 just apply some bad smell rule generated from another project.
 Rather, they say, decisions about refactoring
 should be based based on a combination of (1)~subjective evaluations and 
 (2)~code metrics.

 Given the nature of human
 expertise, developers will offer many  strong opinions about what requires changing.  That is, we   anticipate no shortage of (1)~subjective
 evaluations from developers.

 As to lessons learned from (2)~code metrics, this paper introduces RANK,
 a tool for sorting proposed changes to static code measures. Unlike
 normal classification algorithms, RANK is a {\em contrast set learner}.
  Contrast set learners do not report {\em what is}. Rather
     they recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code refactoring.

RANK's concept of operation is that developers {\em propose}
numerous refactorings and RANK {\em disposes} of all those with
no support in history data.
Experiments with RANK explain the pessimism of Mantyla, Yamashita, 
 and Sj\"oberg et al. As they noted, sometimes, pre-defined bad
 smells are irrelevant to reducing the risk of defects in a project.
 However, when bad smells are learned from the current project,
 it is possible to recognize what are useful, and useless, local
 changes (where ``useful'' is defined in terms of reducing the
 predicted number of defects).
  Hence,   we  can endorse
 the use of bad smells to guide refactorings, but
 only when tools like RANK can  tune the definitions of
 ``good'' or ``bad'' refactorings using current project data.

  The contributions of this paper are:
  \begin{itemize}
      \item A
  \end{itemize}

The rest of this paper is structured as follows. XXX

Before beginning, we make the obvious aside
that defect predictors are not the {\em only} way to assess
bad smells. Some researchers report that code reviews and debating potential refactorings are useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. That said,
reducing the risk of errors was
Fowler's original motivation for introducing refactoring. Also, using
defect predictors,
it is possible to operationalize this test for $R_i \succ R_j$
while it might be harder to otherwise assess refactorings.
Hence, 
this paper assumes that {\em better} refactorings and
bettering at {\em reducing} the chances
that code will contain defects.

@@

A frequently asked question about this work is 
why build tools like RANK to second-guess and critique developer intuitions?
Our reply is that  recent studies point to a  widespread malaise in software engineering:
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
strong views on many issues, including bad smells, but those views may be wrong for the current
project.

To be fair, software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer several
specific studies
showing that   widely-held views in SE that are   now questionable, given recent results. 
The first three come from the general SE literature while the rest
relate specifically to bad smells:
\begin{enumerate} 
\item
Nagappan et al. analyzed~\cite{mei15} 
how   {\tt goto}  was used  in 11,000 Github repositories. 
They found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. 
We found no evidence for such large   phase delay effect in modern
software\footnote{171 software projects shepherded
by the SEI, 2006-2014} possibly this has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices). 
Whatever the reason, the main point here is that phase delay remains a strongly
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

\input{tex/refactor}

 If widely-held beliefs in SE can be inaccurate and misleading,
 it is timely to check the utility of bad smells.
 This check is important since
 the concept of bad smells is highly pervasive in the SE literature\footnote{.
As of  Feburary 2016, ACM Digital Portal reports 
that Fowler's 1999 text has 6434 references;
while a search  on "bad smell"+refactoring yields 824 citations in
the last decade at learning research
venues (ICSE, FSE, etc).}. Accordingly,
\fig{smells} 
lists the commonly mentioned bad smells, in the first column.
The other
columns show other work that comments on which bad smells matter most.
A  blank cell indicates where the other work has chosen to ignore
of the Fowler bad smells as unimportant. 
Note that most of the cells are blank.
For example, the Atwood'06~\cite{Atwood06} study takes the time to comment on   
all the Fowler bad smells.
However, all the  other  studies ignore the majority of the Fowler bad smells.

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells:
\begin{itemize}
\item The Yamashita’13~\cite{Yamashita2013} survey comments on the relative importance
    of 14 code smells-- only eight of which are mentioned in
    the left-hand-column.
  \item  
   The {\em developer survey}   (in the right-hand-side column}
   shows the opinions we collected from a dozen developers from a Washington
    D.C. web-tools development organization
     Amongst those group, there was there was  some
    consensus on  the priority of fixing which  bad smells   
    (see the annotations VH=very high,
    H=high, L=low, VL=very low, and ``?''= no consensus).  
\end{itemize} 
    One interesting point of agreement of the  Yamashita’13 study and the 
    developer survey is the agreement on which  bad smells
    were high priority:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey.
        \item Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \ei
 Note that this agreement also means that the
  Yamashita’13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring refactoring.

In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be flawed.
Hence it is important to double check their beliefs in, for example,
bad smells. Tools like HISTORY perform that check, using models learned
from the historical record.

\input{tex/jur_char}

\subsection{Correlation, Causation, and Critique}

We propose HISTORY as way to critique (and possibly dismiss) proposed
refactorings. This section discusses the value of that approach.

HISTORY is a statistical method for finding ways to change code
such that the new code has a lower chance of having defects. 
To make these recommendations,
HISTORY  inputs frequency counts of how often a static
code attribute is seen in defective or non-defective modules.
Hence, its recommendations about what to change may be wrong
due to the famous correlation vs causation problem.
Statistical methods that use frequency counts can only see how often
certain events occur together, but not why that happens. Hence,
they may conclude incorrectly that with two co-located situations,
one might cause the other. This is not necessarily true. For
example, the number of storks and the number of births
in Denmark is not causally
connected. Nevertheless, both may be correlated if certain
migratory patterns of storks coincide with certain sociological factors.

Another reason that HISTORY's recommendations might be incorrect
is the presence of hidden constraints that are not apparent in
its input data.

@@

It turns out that ranking codiiiiie smells, and ignoring the low priority
ones, is a pressing problem. As shown in our {\em Related Work} section,
different developers and  books and tools stress the importance of different code smells. How to assess these different claims? How to avoid
needless refactoring? How to check if a code smell found at another site
is relevant to this site?

To answer all these questions, we propose NOSE, a tool that checks
if cpde smells really smell 
generates candidate refactorings to reduce the chances of software
defects. XTREE is a back-end to a decision tree learner that
reflects on how different branches in that tree predict for
defective or non-defective code modules. Since this tree
is built from static code attributes, the differences between
those brnachesy reflecting on the
{\em delta} between those branches, it is

Fowler claims   removing   code smells   by
\begin{quote}
...applying a series of small behavior-preserving transformations, each 
of which seem ``too small to be worth doing''. 
The  effect of   these transformations is quite significant. By doing them in small steps you {\em reduce the risk of introducing errors}'' (our italics).
\end{quote}
The italics in the last paragraph offer one method for indicating if bad smell-based
refactoring was successful. There are many automatic
methods for learning   defect predictors from static
code features~\cite{peters15,he13}.  We can use those
predictors to rank $N$ proposed refactorings $R_1,R_2,...R_i$
refactorings. Specifically, we  say that refactoring $i$ is preferred to $j$
(denoted $R_i \succ R_j$) if the defect predictors predict fewer errors
after applying $R_i$ than $R_j$.

The ability to rank, and perhaps   reject, a proposed refactoring is
an important contribution, given the current level of interest
in bad smell-based refactoring.

Nevertheless,
certain results raise doubts about the value of   code smells for reducing the risk of defects in software.
